Constructing a universe for the setoid model

### 2020-07-04 paper.tex
Adapt section Indcutve-recursive

In this section we give a first definition of the setoid universe, as a direct
generalization of the simple inductive-recursive definition just shown. We only
consider a very small universe with unit type $\mathds{1}$ and $\Pi$ for
simplicity; a more realistic universe can be found in the Agda formalization,
and includes more type formers.

@@

We give an inductive definition of $\setoidU$, mutually with a recursive
definition of everything else. Below we only show the more interesting
definitions of $\setoidU$ and $\El$:

\begin{minipage}{0.5\textwidth}
\begin{align*}
  & \textsf{data}\ \setoidU : \Set_1 \\
  & \quad \textsf{one} : \setoidU \\
  & \quad \textsf{pi}
  :  (A : \setoidU) (B : \El\ A \to \setoidU)\\
  & \ \qquad \to (\{x\ x' : \El\ A\} \to \textsf{refl}_\setoidU\ A \vdash x \sim_\El x'\\
  & \qquad\quad \to B\ x \sim_\setoidU B\ x')
  \to \setoidU
\end{align*}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{align*}
  & \El\ \textsf{one} :\equiv \textbf{1} \\
  & \El\ (\textsf{pi}\ A\ B\ h) :\equiv \\
  & \qquad \Sigma\ (f : (a : \El\ A) \to \El\ (B\ a)) \\
  & \qquad \quad
  (\forall\{x\ x'\}(p : \eqel{\reflu\ A}{x}{x'}) \\
  & \qquad \quad\to \eqel{h\ p}{f\ x}{f\ x'})
\end{align*}
\end{minipage}

@@

We can show that $\Uty$ is closed under $\Pi$ types and unit types, including
$\beta$ and $\eta$ laws. Moreover, all the necessary equations (including
substitution laws) hold definitionally in the metatheory.

>>>>

In this section we give a first definition of the setoid universe, as a direct
generalization of the simple inductive-recursive definition just shown. We only
consider a very small universe with bool type $\mathbbm{2}$ and $\Pi$ for
simplicity; a more realistic universe can be found in the Agda formalization,
and includes more type formers.

@@

We give an inductive definition of $\setoidU$, mutually with a recursive definition
of the 4 functions
$\_\sim_\setoidU\_$, $\textsf{refl}_\setoidU$, $\El$ and $\_\vdash\_\sim_\El\_$.
The other functions are then recursively defined:
\begin{itemize}
\item $\textsf{refl}_\El$ alone,
\item $\textsf{sym}_\setoidU$ and $\textsf{sym}_\El$ mutually,
\item $\textsf{trans}_\setoidU$, $\textsf{trans}_\El$, $\coeel$ and $\cohel$ mutually.
\end{itemize}
The whole construction is quite long, below we only show the more interesting
definitions of $\setoidU$ and $\El$:

\begin{minipage}{0.5\textwidth}
\begin{align*}
  & \textsf{data}\ \setoidU : \Set_1 \\
  & \quad \textsf{bool} : \setoidU \\
  & \quad \textsf{pi}
  :  (A : \setoidU) (B : \El\ A \to \setoidU)\\
  & \ \qquad \to (\{x\ x' : \El\ A\} \to \textsf{refl}_\setoidU\ A \vdash x \sim_\El x'\\
  & \qquad\quad \to B\ x \sim_\setoidU B\ x')
  \to \setoidU
\end{align*}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{align*}
  & \El\ \textsf{bool} :\equiv \mathbbm{2} \\
  & \El\ (\textsf{pi}\ A\ B\ h) :\equiv \\
  & \qquad \Sigma\ (f : (a : \El\ A) \to \El\ (B\ a)) \\
  & \qquad \quad
  (\forall\{x\ x'\}(p : \eqel{\reflu\ A}{x}{x'}) \\
  & \qquad \quad\to \eqel{h\ p}{f\ x}{f\ x'})
\end{align*}
\end{minipage}

@@

We can show that $\Uty$ is closed under $\Pi$ types and bool type, including
$\beta$ and $\eta$ laws. Moreover, all the necessary equations (including
substitution laws) hold definitionally in the metatheory.
The universe can be closed under more constructions if more codes are added to
$\setoidU$.
